04_DESeq_sexes

### 2019-08-28 analysis/04_DESeq2_sexes.Rmd
pitutiary lda, now a fucntion)

library(tidyverse)
library(DESeq2)
library(cowplot)
library(RColorBrewer)
library(pheatmap)
library(kableExtra)
library(viridis)
library(forcats)
library("wesanderson")

@@

# set levels
c.colData$treatment <- factor(c.colData$treatment, levels = 
                              c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9"))
levels(c.colData$treatment)
c.colData$sextissue <- as.factor(paste(c.colData$sex, c.colData$tissue, sep = "_"))
summary(c.colData[c(7,3,4,5,8)])
```

## Run DESeq on all subsets of the data

@@

>>>>

library(tidyverse)
library(DESeq2)
library(cowplot)
library(RColorBrewer)
library(pheatmap)
library(kableExtra)
library(viridis)
library(forcats)
library("wesanderson")
library(caret) # LDA analysis
library(MASS) # LDA analysis

@@

# set levels
c.colData$treatment <- factor(c.colData$treatment, levels = 
                              c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9"))
levels(c.colData$treatment)
c.colData$sextissue <- as.factor(paste(c.colData$sex, c.colData$tissue, sep = "_"))

summary(c.colData[c(7,3,4,5,8)])

c.colData <- c.colData %>%
    mutate(hypothesis = fct_recode(treatment,
                            "anticipation" = "control",
                            "anticipation" = "bldg",
                            "incubation" = "lay",
                            "incubation" = "inc.d3",
                            "incubation" = "inc.d9",
                            "incubation" = "inc.d17",
                            "hatchling.care" = "hatch",
                            "hatchling.care" = "n5",
                            "hatchling.care" = "n9"))

```

@@

## Linear discriminant analysis (LDA)

http://www.sthda.com/english/articles/36-classification-methods-essentials/146-discriminant-analysis-essentials-in-r/

```{r LDA}
theme_set(theme_classic())

LDAanalysis <- function(mydds){
  # data prep
  vsd <- vst(mydds, blind=FALSE) # variance stabilized 
  tvsd <- as.data.frame(t(assay(vsd)))   # get vsd counts and transform

  tvsd <- tvsd[, sample(ncol(tvsd), 20)] # select 20 random genes
  tvsd$V1 <- row.names(tvsd)             # prep for joining

  colDataPit <- subsetcolData2(c.colData, c("female_pituitary", "male_pituitary"))

  mydata <- left_join(tvsd, colDataPit, by = "V1")  # merge counts and variables

  # Split the data into training (80%) and test set (20%)
  set.seed(175)

  training.samples <- mydata$treatment %>%
    createDataPartition(p = 0.8, list = FALSE)
  train.data <- mydata[training.samples, ]
  test.data <- mydata[-training.samples, ]

  # Normalize the data. Categorical variables are automatically ignored.
  # Estimate preprocessing parameters
  preproc.param <- train.data %>% 
    preProcess(method = c("center", "scale"))

  # Transform the data using the estimated parameters
  train.transformed <- preproc.param %>% predict(train.data)
  test.transformed <- preproc.param %>% predict(test.data)

  # LDA analysis

  # Fit the model
  print(model <- lda(treatment~ sex, data = train.transformed))
  # Make predictions
  print(predictions <- model %>% predict(test.transformed))
# Model accuracy
  print(mean(predictions$class==test.transformed$treatment))

  # results
  print(model)

  # make predictions
  predictions <- model %>% predict(test.transformed)

  # Predicted classes
  print(head(predictions$class, 6))
  # Predicted probabilities of class memebership.
  print(head(predictions$posterior, 6) )
  # Linear discriminants
  print(head(predictions$x, 3) )

  # LDA plot using ggplot2 
  lda.data <- cbind(train.transformed, predict(model)$x)
  p<-ggplot(data=lda.data, aes(x=treatment, y=LD1, fill = treatment)) +
    geom_bar( stat = "summary", fun.y = "mean")
  p

  # model accuracy 
  print(mean(predictions$class==test.transformed$treatment))
}

LDAanalysis(dds.pituitary)
```
